---
title: 'Exploratory Data Analysis on los_angeles_12.csv'
author: "USC Machine Learning Team"
date: "11/2/2017"
output: pdf_document
header-includes:
  - \usepackage{color}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Loading

```{r data loading, echo=TRUE}
# load data file
if (!exists("LA12.raw")) {
    LA12.raw <- read.csv("./los_angeles_12.csv")
}
print(nrow(LA12.raw))
```

# Data Cleaning & Exploratory Analysis

Data Cleaning & Exploratory Analysis are iterative, starting at next page.

\pagebreak 

## Epoch 1: Display of Vertical Acceleration with Natural Gravity

**Date**: the time stamp of a data point \newline 
**Latitude**: \newline 
**Longitude**: \newline 
**Speed**: traveling speed of the vehicle \newline 
**speedbump**: whether this data point is a speedbump \newline
**forw_accel**: forward acceleration (front and back) of the vehicle \newline 
**hori_accel**: horizontal acceleration (left and right) of the vehicle \newline 
**vert_accel_G**: vertical acceleration (up and down) of the vehicle with natural gravity (-1G) \newline

```{r data cleaning 1, echo=TRUE}
LA12.valid = LA12.raw[, c("Date", "Latitude", "Longitude", "Speed", "speedbump")]
# validate DateTime format
LA12.valid$Date <- as.POSIXct(LA12.valid$Date, format="%Y-%m-%d %H:%M:%OS")
# specify orientation of accelration
LA12.valid$forw_accel = LA12.raw$X
LA12.valid$hori_accel = LA12.raw$Y
LA12.valid$vert_accel_G = LA12.raw$Z
```

```{r exploratory data analysis 1, echo=FALSE}
require(ggplot2)
LA12.vert_accel_G.plot <- ggplot(LA12.valid, aes(x = Date))
LA12.vert_accel_G.plot <- LA12.vert_accel_G.plot + geom_line(aes(y = vert_accel_G, colour = "Vert. Accel. with -1G"))
LA12.vert_accel_G.plot <- LA12.vert_accel_G.plot + scale_colour_manual(values = c("darkorchid4"))
LA12.vert_accel_G.plot <- LA12.vert_accel_G.plot + xlab("Time") 
LA12.vert_accel_G.plot <- LA12.vert_accel_G.plot + ylab("Vertical Acceleration (G-Force)") 
LA12.vert_accel_G.plot <- LA12.vert_accel_G.plot + ggtitle("Time-Indexed Vertical Acceleration with Natural Gravity")
LA12.vert_accel_G.plot <- LA12.vert_accel_G.plot + theme(legend.position = c(0.8, 0.9))
for (i in 1:nrow(LA12.valid)) {
    if (LA12.valid$speedbump[i] == "yes") {
        LA12.vert_accel_G.plot <- LA12.vert_accel_G.plot + 
            annotate("pointrange", x = LA12.valid$Date[i], y = -1, ymin = -1, ymax = -1, colour = "red", size = 0.25)
    }
}
print(LA12.vert_accel_G.plot)
```

Note: Speed bumps are labeled as $\color{red}red$ points on the graph.

\pagebreak

## Epoch 2: Display of Vertical Acceleration without Natural Gravity

**vert_accel**: vertical acceleration (up and down) of the vehicle without natural gravity \newline

```{r data cleaning 2, echo=TRUE}
# remove natural gravity in vertical acceleration
LA12.valid$vert_accel = LA12.valid$vert_accel_G + 1
```

```{r exploratory data analysis 2, echo=FALSE}
require(ggplot2)
LA12.vert_accel.plot <- ggplot(LA12.valid, aes(x = Date))
LA12.vert_accel.plot <- LA12.vert_accel.plot + geom_line(aes(y = vert_accel, colour = "Vertical Acceleration"))
LA12.vert_accel.plot <- LA12.vert_accel.plot + geom_line(aes(y = Speed / 10, colour = "Traveling Speed"))
LA12.vert_accel.plot <- LA12.vert_accel.plot + scale_y_continuous(sec.axis = sec_axis(~.*10, name = "Traveling Speed (m/s)"))
LA12.vert_accel.plot <- LA12.vert_accel.plot + scale_colour_manual(values = c("deepskyblue", "darkorchid4"))
LA12.vert_accel.plot <- LA12.vert_accel.plot + xlab("Time") 
LA12.vert_accel.plot <- LA12.vert_accel.plot + ylab("Vertical Acceleration (G-Force)") 
LA12.vert_accel.plot <- LA12.vert_accel.plot + ggtitle("Time-Indexed Vertical Acceleration without Natural Gravity")
LA12.vert_accel.plot <- LA12.vert_accel.plot + theme(legend.position = c(0.8, 0.9))
for (i in 1:nrow(LA12.valid)) {
    if (LA12.valid$speedbump[i] == "yes") {
        LA12.vert_accel.plot <- LA12.vert_accel.plot + 
            annotate("pointrange", x = LA12.valid$Date[i], y = 0, ymin = 0, ymax = 0, colour = "red", size = 0.25)
    }
}
print(LA12.vert_accel.plot)
```

Note: Speed bumps are labeled as $\color{red}red$ points on the graph.

\pagebreak

## Epoch 3: Display of Ratio between Vertical Acceleration (w/o Natural G) and Speed

**vert_accel_ratio_speed**: the ratio between vertical acceleration (without natural gravity) and traveling speed \newline

```{r data cleaning 3, echo=TRUE}
# calculate ratio between vertical acceleration and speed
for (i in 1:nrow(LA12.valid)) {
    if (LA12.valid$Speed[i] == 0) {
        LA12.valid$vert_accel_ratio_speed[i] = 0
    }
    else {
        LA12.valid$vert_accel_ratio_speed[i] = LA12.valid$vert_accel[i] / LA12.valid$Speed[i]
    }
}
```

```{r exploratory data analysis 3, echo=FALSE}
require(ggplot2)
LA12.vert_accel_ratio_speed.plot <- ggplot(LA12.valid, aes(x = Date))
LA12.vert_accel_ratio_speed.plot <- LA12.vert_accel_ratio_speed.plot + geom_line(aes(y = vert_accel_ratio_speed, colour = "Ratio of Vert. Accel. and Speed"))
LA12.vert_accel_ratio_speed.plot <- LA12.vert_accel_ratio_speed.plot + geom_line(aes(y = Speed / 100, colour = "Traveling Speed"))
LA12.vert_accel_ratio_speed.plot <- LA12.vert_accel_ratio_speed.plot + scale_y_continuous(sec.axis = sec_axis(~.*100, name = "Traveling Speed (m/s)"))
LA12.vert_accel_ratio_speed.plot <- LA12.vert_accel_ratio_speed.plot + scale_colour_manual(values = c("darkorchid4", "deepskyblue"))
LA12.vert_accel_ratio_speed.plot <- LA12.vert_accel_ratio_speed.plot + xlab("Time") 
LA12.vert_accel_ratio_speed.plot <- LA12.vert_accel_ratio_speed.plot + ylab("Ratio of Vert. Accel. and Speed") 
LA12.vert_accel_ratio_speed.plot <- LA12.vert_accel_ratio_speed.plot + ggtitle("Time-Indexed Ratio of Vertical Acceleration and Speed")
LA12.vert_accel_ratio_speed.plot <- LA12.vert_accel_ratio_speed.plot + theme(legend.position = c(0.8, 0.9))
for (i in 1:nrow(LA12.valid)) {
    if (LA12.valid$speedbump[i] == "yes") {
        LA12.vert_accel_ratio_speed.plot <- LA12.vert_accel_ratio_speed.plot + 
            annotate("pointrange", x = LA12.valid$Date[i], y = 0, ymin = 0, ymax = 0, colour = "red", size = 0.25)
    }
}
print(LA12.vert_accel_ratio_speed.plot)
```

Note: Speed bumps are labeled as $\color{red}red$ points on the graph. \newline

\pagebreak

## Epoch 4: Display of Ratio between Vertical Acceleration (w/o Natural G) and Speed

**sq_vert_accel_ratio_speed**: the ratio between squared vertical acceleration (without natual gravity) and traveling speed \newline

```{r data cleaning 4, echo=TRUE}
# calculate ratio between vertical acceleration and speed
for (i in 1:nrow(LA12.valid)) {
    if (LA12.valid$Speed[i] == 0) {
        LA12.valid$sq_vert_accel_ratio_speed[i] = 0
    }
    else {
        LA12.valid$sq_vert_accel_ratio_speed[i] = (LA12.valid$vert_accel[i] * LA12.valid$vert_accel[i]) / LA12.valid$Speed[i]
    }
}
```

```{r exploratory data analysis 4, echo=FALSE}
require(ggplot2)
LA12.sq_vert_accel_ratio_speed.plot <- ggplot(LA12.valid, aes(x = Date))
LA12.sq_vert_accel_ratio_speed.plot <- LA12.sq_vert_accel_ratio_speed.plot + geom_line(aes(y = sq_vert_accel_ratio_speed, colour = "Ratio of Sq. Vert. Accel. and Speed"))
LA12.sq_vert_accel_ratio_speed.plot <- LA12.sq_vert_accel_ratio_speed.plot + geom_line(aes(y = Speed / 50, colour = "Traveling Speed"))
LA12.sq_vert_accel_ratio_speed.plot <- LA12.sq_vert_accel_ratio_speed.plot + scale_y_continuous(sec.axis = sec_axis(~.*50, name = "Traveling Speed (m/s)"))
LA12.sq_vert_accel_ratio_speed.plot <- LA12.sq_vert_accel_ratio_speed.plot + scale_colour_manual(values = c("darkorchid4", "deepskyblue"))
LA12.sq_vert_accel_ratio_speed.plot <- LA12.sq_vert_accel_ratio_speed.plot + xlab("Time") 
LA12.sq_vert_accel_ratio_speed.plot <- LA12.sq_vert_accel_ratio_speed.plot + ylab("Ratio of Sq. Vert. Accel. and Speed") 
LA12.sq_vert_accel_ratio_speed.plot <- LA12.sq_vert_accel_ratio_speed.plot + ggtitle("Time-Indexed Ratio of Squared Vertical Acceleration and Speed")
LA12.sq_vert_accel_ratio_speed.plot <- LA12.sq_vert_accel_ratio_speed.plot + theme(legend.position = c(0.8, 0.9))
for (i in 1:nrow(LA12.valid)) {
    if (LA12.valid$speedbump[i] == "yes") {
        LA12.sq_vert_accel_ratio_speed.plot <- LA12.sq_vert_accel_ratio_speed.plot + 
            annotate("pointrange", x = LA12.valid$Date[i], y = 0, ymin = 0, ymax = 0, colour = "red", size = 0.25)
    }
}
print(LA12.sq_vert_accel_ratio_speed.plot)
```

Note: Speed bumps are labeled as $\color{red}red$ points on the graph.

\pagebreak

## Epoch 5: Display of Vertical Jolt (Incremental Change of Vert. Accel.)

**vert_jolt**: vertical jolt of the vehicle (incremental change of vertical acceleration) \newline

```{r data cleaning 5, echo=TRUE}
# calculate vertical jolt
for (i in 1:nrow(LA12.valid)) {
    if (i == 1) {
        LA12.valid$vert_jolt[i] = 0
    }
    else {
        LA12.valid$vert_jolt[i] = LA12.valid$vert_accel[i] - LA12.valid$vert_accel[i - 1]
    }
}
```

```{r exploratory data analysis 5, echo=FALSE}
require(ggplot2)
LA12.vert_jolt.plot <- ggplot(LA12.valid, aes(x = Date))
LA12.vert_jolt.plot <- LA12.vert_jolt.plot + geom_line(aes(y = vert_jolt, colour = "Vertical Jolt"))
LA12.vert_jolt.plot <- LA12.vert_jolt.plot + geom_line(aes(y = Speed / 10, colour = "Traveling Speed"))
LA12.vert_jolt.plot <- LA12.vert_jolt.plot + scale_y_continuous(sec.axis = sec_axis(~.*10, name = "Traveling Speed (m/s)"))
LA12.vert_jolt.plot <- LA12.vert_jolt.plot + scale_colour_manual(values = c("deepskyblue", "darkorchid4"))
LA12.vert_jolt.plot <- LA12.vert_jolt.plot + xlab("Time") 
LA12.vert_jolt.plot <- LA12.vert_jolt.plot + ylab("Vertical Jolt (G-Force)") 
LA12.vert_jolt.plot <- LA12.vert_jolt.plot + ggtitle("Time-Indexed Vertical Jolt (Incremental)")
LA12.vert_jolt.plot <- LA12.vert_jolt.plot + theme(legend.position = c(0.8, 0.9))
for (i in 1:nrow(LA12.valid)) {
    if (LA12.valid$speedbump[i] == "yes") {
        LA12.vert_jolt.plot <- LA12.vert_jolt.plot + 
            annotate("pointrange", x = LA12.valid$Date[i], y = 0, ymin = 0, ymax = 0, colour = "red", size = 0.25)
    }
}
print(LA12.vert_jolt.plot)
```

Note: Speed bumps are labeled as $\color{red}red$ points on the graph.

\pagebreak

## Epoch 6: Display of Ratio between Vertical Jolt and Speed

**vert_jolt_ratio_speed**: the ratio between vertical jolt and traveling speed \newline

```{r data cleaning 6, echo=TRUE}
# calculate ratio between vertical jolt and speed
for (i in 1:nrow(LA12.valid)) {
    if (LA12.valid$Speed[i] == 0) {
        LA12.valid$vert_jolt_ratio_speed[i] = 0
    }
    else {
        LA12.valid$vert_jolt_ratio_speed[i] = LA12.valid$vert_jolt[i] / LA12.valid$Speed[i]
    }
}
```

```{r exploratory data analysis 6, echo=FALSE}
require(ggplot2)
LA12.vert_jolt_ratio_speed.plot <- ggplot(LA12.valid, aes(x = Date))
LA12.vert_jolt_ratio_speed.plot <- LA12.vert_jolt_ratio_speed.plot + geom_line(aes(y = vert_jolt_ratio_speed, colour = "Ratio of Vert. Jolt and Speed"))
LA12.vert_jolt_ratio_speed.plot <- LA12.vert_jolt_ratio_speed.plot + geom_line(aes(y = Speed / 100, colour = "Traveling Speed"))
LA12.vert_jolt_ratio_speed.plot <- LA12.vert_jolt_ratio_speed.plot + scale_y_continuous(sec.axis = sec_axis(~.*100, name = "Traveling Speed (m/s)"))
LA12.vert_jolt_ratio_speed.plot <- LA12.vert_jolt_ratio_speed.plot + scale_colour_manual(values = c("darkorchid4", "deepskyblue"))
LA12.vert_jolt_ratio_speed.plot <- LA12.vert_jolt_ratio_speed.plot + xlab("Time") 
LA12.vert_jolt_ratio_speed.plot <- LA12.vert_jolt_ratio_speed.plot + ylab("Ratio of Vert. Jolt and Speed") 
LA12.vert_jolt_ratio_speed.plot <- LA12.vert_jolt_ratio_speed.plot + ggtitle("Time-Indexed Ratio of Vertical Jolt and Speed")
LA12.vert_jolt_ratio_speed.plot <- LA12.vert_jolt_ratio_speed.plot + theme(legend.position = c(0.8, 0.9))
for (i in 1:nrow(LA12.valid)) {
    if (LA12.valid$speedbump[i] == "yes") {
        LA12.vert_jolt_ratio_speed.plot <- LA12.vert_jolt_ratio_speed.plot + 
            annotate("pointrange", x = LA12.valid$Date[i], y = 0, ymin = 0, ymax = 0, colour = "red", size = 0.25)
    }
}
print(LA12.vert_jolt_ratio_speed.plot)
```

Note: Speed bumps are labeled as $\color{red}red$ points on the graph.

\pagebreak

## Epoch 7: Display of Ratio between Squared Vertical Jolt and Speed

**sq_vert_jolt_ratio_speed**: the ratio between squared vertical jolt and traveling speed \newline

```{r data cleaning 7, echo=TRUE}
# calculate ratio between vertical jolt and speed
for (i in 1:nrow(LA12.valid)) {
    if (LA12.valid$Speed[i] == 0) {
        LA12.valid$sq_vert_jolt_ratio_speed[i] = 0
    }
    else {
        LA12.valid$sq_vert_jolt_ratio_speed[i] = (LA12.valid$vert_jolt[i] * LA12.valid$vert_jolt[i]) / LA12.valid$Speed[i]
    }
}
```

```{r exploratory data analysis 7, echo=FALSE}
require(ggplot2)
LA12.sq_vert_jolt_ratio_speed.plot <- ggplot(LA12.valid, aes(x = Date))
LA12.sq_vert_jolt_ratio_speed.plot <- LA12.sq_vert_jolt_ratio_speed.plot + geom_line(aes(y = sq_vert_jolt_ratio_speed, colour = "Ratio of Sq. Vert. Jolt and Speed"))
LA12.sq_vert_jolt_ratio_speed.plot <- LA12.sq_vert_jolt_ratio_speed.plot + geom_line(aes(y = Speed / 50, colour = "Traveling Speed"))
LA12.sq_vert_jolt_ratio_speed.plot <- LA12.sq_vert_jolt_ratio_speed.plot + scale_y_continuous(sec.axis = sec_axis(~.*50, name = "Traveling Speed (m/s)"))
LA12.sq_vert_jolt_ratio_speed.plot <- LA12.sq_vert_jolt_ratio_speed.plot + scale_colour_manual(values = c("darkorchid4", "deepskyblue"))
LA12.sq_vert_jolt_ratio_speed.plot <- LA12.sq_vert_jolt_ratio_speed.plot + xlab("Time") 
LA12.sq_vert_jolt_ratio_speed.plot <- LA12.sq_vert_jolt_ratio_speed.plot + ylab("Ratio of Sq. Vert. Jolt and Speed") 
LA12.sq_vert_jolt_ratio_speed.plot <- LA12.sq_vert_jolt_ratio_speed.plot + ggtitle("Time-Indexed Ratio of Squared Vertical Jolt and Speed")
LA12.sq_vert_jolt_ratio_speed.plot <- LA12.sq_vert_jolt_ratio_speed.plot + theme(legend.position = c(0.8, 0.9))
for (i in 1:nrow(LA12.valid)) {
    if (LA12.valid$speedbump[i] == "yes") {
        LA12.sq_vert_jolt_ratio_speed.plot <- LA12.sq_vert_jolt_ratio_speed.plot + 
            annotate("pointrange", x = LA12.valid$Date[i], y = 0, ymin = 0, ymax = 0, colour = "red", size = 0.25)
    }
}
print(LA12.sq_vert_jolt_ratio_speed.plot)
```

Note: Speed bumps are labeled as $\color{red}red$ points on the graph.

\pagebreak

## Epoch 8: Display of Sliding-Window Statistics of Vertical Jolt

**vert_jolt_mean**: 5-sliding-window mean of vertical jolt \newline 
**vert_jolt_sd**: 5-sliding-window standard deviation of vertical jolt \newline 
**vert_jolt_min**: 5-sliding-window minimum of vertical jolt \newline 
**vert_jolt_max**: 5-sliding-window maximum of vertical jolt \newline 
**vert_jolt_range**: 5-sliding-window range of vertical jolt \newline 

```{r data cleaning 8, echo=TRUE}
# calculate 5-sliding-window mean of vertical jolt
for (i in 3:nrow(LA12.valid)-2) {
    LA12.valid$vert_jolt_mean[i] = mean(c(LA12.valid$vert_jolt[i-2], 
                                        LA12.valid$vert_jolt[i-1], 
                                        LA12.valid$vert_jolt[i], 
                                        LA12.valid$vert_jolt[i+1], 
                                        LA12.valid$vert_jolt[i+2]))
}
LA12.valid$vert_jolt_mean[1] = 0
LA12.valid$vert_jolt_mean[2] = 0
LA12.valid$vert_jolt_mean[nrow(LA12.valid) - 1] = 0
LA12.valid$vert_jolt_mean[nrow(LA12.valid)] = 0
# calculate 5-sliding-window standard deviation of vertical jolt
for (i in 3:nrow(LA12.valid)-2) {
    LA12.valid$vert_jolt_sd[i] = sd(c(LA12.valid$vert_jolt[i-2], 
                                        LA12.valid$vert_jolt[i-1], 
                                        LA12.valid$vert_jolt[i], 
                                        LA12.valid$vert_jolt[i+1], 
                                        LA12.valid$vert_jolt[i+2]))
}
LA12.valid$vert_jolt_sd[1] = 0
LA12.valid$vert_jolt_sd[2] = 0
LA12.valid$vert_jolt_sd[nrow(LA12.valid) - 1] = 0
LA12.valid$vert_jolt_sd[nrow(LA12.valid)] = 0
# calculate 5-sliding-window minimum of vertical jolt
for (i in 3:nrow(LA12.valid)-2) {
    LA12.valid$vert_jolt_min[i] = min(c(LA12.valid$vert_jolt[i-2], 
                                        LA12.valid$vert_jolt[i-1], 
                                        LA12.valid$vert_jolt[i], 
                                        LA12.valid$vert_jolt[i+1], 
                                        LA12.valid$vert_jolt[i+2]))
}
LA12.valid$vert_jolt_min[1] = 0
LA12.valid$vert_jolt_min[2] = 0
LA12.valid$vert_jolt_min[nrow(LA12.valid) - 1] = 0
LA12.valid$vert_jolt_min[nrow(LA12.valid)] = 0
# calculate 5-sliding-window maxium of vertical jolt
for (i in 3:nrow(LA12.valid)-2) {
    LA12.valid$vert_jolt_max[i] = max(c(LA12.valid$vert_jolt[i-2], 
                                        LA12.valid$vert_jolt[i-1], 
                                        LA12.valid$vert_jolt[i], 
                                        LA12.valid$vert_jolt[i+1], 
                                        LA12.valid$vert_jolt[i+2]))
}
LA12.valid$vert_jolt_max[1] = 0
LA12.valid$vert_jolt_max[2] = 0
LA12.valid$vert_jolt_max[nrow(LA12.valid) - 1] = 0
LA12.valid$vert_jolt_max[nrow(LA12.valid)] = 0
# calculate 5-sliding-window range of vertical jolt
LA12.valid$vert_jolt_range = LA12.valid$vert_jolt_max - LA12.valid$vert_jolt_min
```

```{r exploratory data analysis 8, echo=FALSE}
require(ggplot2)
LA12.vert_jolt_5.plot <- ggplot(LA12.valid, aes(x = Date))
LA12.vert_jolt_5.plot <- LA12.vert_jolt_5.plot + geom_line(aes(y = vert_jolt_range, colour = "Range (sky blue)"))
LA12.vert_jolt_5.plot <- LA12.vert_jolt_5.plot + geom_line(aes(y = vert_jolt_mean, colour = "Mean (dark blue)"))
LA12.vert_jolt_5.plot <- LA12.vert_jolt_5.plot + geom_line(aes(y = vert_jolt_sd, colour = "Standard Deviation (orange)"))
LA12.vert_jolt_5.plot <- LA12.vert_jolt_5.plot + geom_line(aes(y = vert_jolt_min, colour = "Minimum (green)"))
LA12.vert_jolt_5.plot <- LA12.vert_jolt_5.plot + geom_line(aes(y = vert_jolt_max, colour = "Maximum (purple)"))
LA12.vert_jolt_5.plot <- LA12.vert_jolt_5.plot + scale_colour_manual(values = c("darkorchid4", "blue4", "chartreuse4", "deepskyblue", "goldenrod3"))
LA12.vert_jolt_5.plot <- LA12.vert_jolt_5.plot + theme(legend.position = c(0.3, 0.8))
LA12.vert_jolt_5.plot <- LA12.vert_jolt_5.plot + xlab("Time") 
LA12.vert_jolt_5.plot <- LA12.vert_jolt_5.plot + ylab("5-Sliding-Window Stats. of Vertical Jolt") 
LA12.vert_jolt_5.plot <- LA12.vert_jolt_5.plot + ggtitle("5-Sliding-Window Statistics of Vertical Jolt")
for (i in 1:nrow(LA12.valid)) {
    if (LA12.valid$speedbump[i] == "yes") {
        LA12.vert_jolt_5.plot <- LA12.vert_jolt_5.plot + 
            annotate("pointrange", x = LA12.valid$Date[i], y = 0, ymin = 0, ymax = 0, colour = "red", size = 0.25)
    }
}
print(LA12.vert_jolt_5.plot)
```

Note: Speed bumps are labeled as $\color{red}red$ points on the graph.

\pagebreak 

## Epoch 9: Comparative Display of Two Most Promising Factors

```{r exploratory data analysis 9, echo=FALSE}
require(ggplot2)
LA12.comparative.plot <- ggplot(LA12.valid, aes(x = Date))
LA12.comparative.plot <- LA12.comparative.plot + geom_line(aes(y = Speed / 50, colour = "Traveling Speed"))
#LA12.comparative.plot <- LA12.comparative.plot + geom_line(aes(y = vert_accel, colour = "Vertical Acceleration"))
#LA12.comparative.plot <- LA12.comparative.plot + geom_line(aes(y = vert_jolt, colour = "Vertical Jolt"))
LA12.comparative.plot <- LA12.comparative.plot + geom_line(aes(y = sq_vert_accel_ratio_speed, colour = "Ratio of Sq. Vert. Accel. & Speed"))
LA12.comparative.plot <- LA12.comparative.plot + geom_line(aes(y = -sq_vert_jolt_ratio_speed, colour = "Ratio of Sq. Vert. Jolt & Speed (-)"))
LA12.comparative.plot <- LA12.comparative.plot + scale_y_continuous(sec.axis = sec_axis(~.*50, name = "Traveling Speed (m/s)"))
LA12.comparative.plot <- LA12.comparative.plot + scale_colour_manual(values = c("darkorchid4", "goldenrod3", "deepskyblue"))
LA12.comparative.plot <- LA12.comparative.plot + xlab("Time") 
LA12.comparative.plot <- LA12.comparative.plot + ylab("Ratios") 
LA12.comparative.plot <- LA12.comparative.plot + ggtitle("Comparison of Two Most Promising Factors")
LA12.comparative.plot <- LA12.comparative.plot + theme(legend.position = c(0.35, 0.2))
for (i in 1:nrow(LA12.valid)) {
    if (LA12.valid$speedbump[i] == "yes") {
        LA12.comparative.plot <- LA12.comparative.plot + 
            annotate("pointrange", x = LA12.valid$Date[i], y = 0, ymin = 0, ymax = 0, colour = "red", size = 0.2)
    }
}
print(LA12.comparative.plot)
```

Note: Ratio of Sq. Vert. Jolt & Speed is negatively filpped for visualization purpose. \newline 
Note: Speed bumps are labeled as $\color{red}red$ points on the graph. \newline

\pagebreak 

# Data Writing

**Date**: the time stamp of a data point \newline 
**Latitude**: \newline 
**Longitude**: \newline 
**Speed**: traveling speed of the vehicle \newline 
**speedbump**: whether this data point is a speedbump \newline 
**forw_accel**: forward acceleration (front and back) of the vehicle \newline 
**hori_accel**: horizontal acceleration (left and right) of the vehicle \newline 
**vert_accel_G**: vertical acceleration (up and down) of the vehicle with natural gravity (-1G) \newline 
**vert_accel**: vertical acceleration (up and down) of the vehicle without natural gravity \newline 
**vert_accel_ratio_speed**: the ratio between vertical acceleration (without natural gravity) and traveling speed \newline 
**sq_vert_accel_ratio_speed**: the ratio between squared vertical acceleration (without natual gravity) and traveling speed \newline 
**vert_jolt**: vertical jolt of the vehicle (incremental change of vertical acceleration) \newline 
**vert_jolt_ratio_speed**: the ratio between vertical jolt and traveling speed \newline 
**sq_vert_jolt_ratio_speed**: the ratio between squared vertical jolt and traveling speed \newline 
**vert_jolt_mean**: 5-sliding-window mean of vertical jolt \newline 
**vert_jolt_sd**: 5-sliding-window standard deviation of vertical jolt \newline 
**vert_jolt_min**: 5-sliding-window minimum of vertical jolt \newline 
**vert_jolt_max**: 5-sliding-window maximum of vertical jolt \newline 
**vert_jolt_range**: 5-sliding-window range of vertical jolt \newline 

```{r data writing, echo=TRUE}
jpeg("./Epoch1_LA12.vert_accel_G.plot.jpeg", width = 1200, height = 750, units = "px")
print(LA12.vert_accel_G.plot)
dev.off()

jpeg("./Epoch2_LA12.vert_accel.plot.jpeg", width = 1200, height = 750, units = "px")
print(LA12.vert_accel.plot)
dev.off()

jpeg("./Epoch3_LA12.vert_accel_ratio_speed.plot.jpeg", width = 1200, height = 750, units = "px")
print(LA12.vert_accel_ratio_speed.plot)
dev.off()

jpeg("./Epoch4_LA12.sq_vert_accel_ratio_speed.plot.jpeg", width = 1200, height = 750, units = "px")
print(LA12.sq_vert_accel_ratio_speed.plot)
dev.off()

jpeg("./Epoch5_LA12.vert_jolt.plot.jpeg", width = 1200, height = 750, units = "px")
print(LA12.vert_jolt.plot)
dev.off()

jpeg("./Epoch6_LA12.vert_jolt_ratio_speed.plot.jpeg", width = 1200, height = 750, units = "px")
print(LA12.vert_jolt_ratio_speed.plot)
dev.off()

jpeg("./Epoch7_LA12.sq_vert_jolt_ratio_speed.plot.jpeg", width = 1200, height = 750, units = "px")
print(LA12.sq_vert_jolt_ratio_speed.plot)
dev.off()

jpeg("./Epoch8_LA12.vert_jolt_5.plot.jpeg", width = 1200, height = 750, units = "px")
print(LA12.vert_jolt_5.plot)
dev.off()

jpeg("./Epoch9_LA12.comparative.plot.jpeg", width = 1200, height = 750, units = "px")
print(LA12.comparative.plot)
dev.off()

write.csv(LA12.valid, "./los_angeles_12_labeled.csv")
```
