---
title: "Exploratory Analysis on Kenya Road Data (Oct. 15)"
author: "USC Machine Learning Team"
date: "11/5/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Loading

```{r data loading, echo=TRUE}
# load data file
if (!exists("Kenya.raw")) {
    Kenya.raw <- read.csv("./kenya_oct_15_data_labeled.csv")
}
print(nrow(Kenya.raw))
Kenya.73 = Kenya.raw[Kenya.raw$Batch_id == 73, c("time", "label", "velocity", "x", "x_raw")]
Kenya.73$time <- as.POSIXct(Kenya.73$time, format="%Y-%m-%d %H:%M:%OS")
Kenya.73$Speed <- Kenya.73$velocity
Kenya.73$vert_accel <- Kenya.73$x_raw
Kenya.73$vert_jolt <- Kenya.73$x
Kenya.73 <- Kenya.73[, c("time", "Speed", "vert_accel", "vert_jolt", "label")]
```

**time**: the time stamp of a data point \newline 
**Speed**: traveling speed of the vehicle \newline 
**vert_accel**: vertical acceleration (up and down) of the vehicle without natural gravity \newline 
**vert_jolt**: vertical jolt of the vehicle (incremental change of vertical acceleration) \newline 

\pagebreak 

# Data Cleaning & Exploratory Analysis

## Epoch 1: Display of Ratio betwwen Vertical Acceleration (w/o Natural G) and Speed

**sq_vert_accel_ratio_speed**: the ratio between squared vertical acceleration (without natual gravity) and traveling speed \newline

```{r data cleaning 1, echo=TRUE}
# calculate ratio between vertical acceleration and speed
for (i in 1:nrow(Kenya.73)) {
    if (Kenya.73$Speed[i] == 0) {
        Kenya.73$sq_vert_accel_ratio_speed[i] = 0
    }
    else {
        Kenya.73$sq_vert_accel_ratio_speed[i] = (Kenya.73$vert_accel[i] * Kenya.73$vert_accel[i]) / Kenya.73$Speed[i]
    }
}
```

```{r exploratory data analysis 1.1, echo=FALSE}
require(ggplot2)
Kenya.73.sq_vert_accel_ratio_speed.plot <- ggplot(Kenya.73[1:500, ], aes(x = time))
Kenya.73.sq_vert_accel_ratio_speed.plot <- Kenya.73.sq_vert_accel_ratio_speed.plot + geom_line(aes(y = sq_vert_accel_ratio_speed))
Kenya.73.sq_vert_accel_ratio_speed.plot <- Kenya.73.sq_vert_accel_ratio_speed.plot + xlab("Time") 
Kenya.73.sq_vert_accel_ratio_speed.plot <- Kenya.73.sq_vert_accel_ratio_speed.plot + ylab("Ratio of Sq. Vert. Accel. and Speed") 
Kenya.73.sq_vert_accel_ratio_speed.plot <- Kenya.73.sq_vert_accel_ratio_speed.plot + ggtitle("Time-Indexed Ratio of Squared Vertical Acceleration and Speed")
for (i in 1:500) {
    if (Kenya.73$label[i] == 1) {
         Kenya.73.sq_vert_accel_ratio_speed.plot <- Kenya.73.sq_vert_accel_ratio_speed.plot + 
             annotate("pointrange", x = Kenya.73$time[i], y = 0, ymin = 0, ymax = 0, colour = "red", size = 0.25)
    }
    if (Kenya.73$label[i] == 2) {
        Kenya.73.sq_vert_accel_ratio_speed.plot <- Kenya.73.sq_vert_accel_ratio_speed.plot + 
             annotate("pointrange", x = Kenya.73$time[i], y = 0, ymin = 0, ymax = 0, colour = "darkorange2", size = 0.25)
    }
}
print(Kenya.73.sq_vert_accel_ratio_speed.plot)
```

\pagebreak

```{r exploratory data analysis 1.2, echo=FALSE}
require(ggplot2)
Kenya.73.sq_vert_accel_ratio_speed.plot <- ggplot(Kenya.73[501:1000, ], aes(x = time))
Kenya.73.sq_vert_accel_ratio_speed.plot <- Kenya.73.sq_vert_accel_ratio_speed.plot + geom_line(aes(y = sq_vert_accel_ratio_speed))
Kenya.73.sq_vert_accel_ratio_speed.plot <- Kenya.73.sq_vert_accel_ratio_speed.plot + xlab("Time") 
Kenya.73.sq_vert_accel_ratio_speed.plot <- Kenya.73.sq_vert_accel_ratio_speed.plot + ylab("Ratio of Sq. Vert. Accel. and Speed") 
Kenya.73.sq_vert_accel_ratio_speed.plot <- Kenya.73.sq_vert_accel_ratio_speed.plot + ggtitle("Time-Indexed Ratio of Squared Vertical Acceleration and Speed")
for (i in 501:1000) {
    if (Kenya.73$label[i] == 1) {
         Kenya.73.sq_vert_accel_ratio_speed.plot <- Kenya.73.sq_vert_accel_ratio_speed.plot + 
             annotate("pointrange", x = Kenya.73$time[i], y = 0, ymin = 0, ymax = 0, colour = "red", size = 0.25)
    }
    if (Kenya.73$label[i] == 2) {
        Kenya.73.sq_vert_accel_ratio_speed.plot <- Kenya.73.sq_vert_accel_ratio_speed.plot + 
             annotate("pointrange", x = Kenya.73$time[i], y = 0, ymin = 0, ymax = 0, colour = "darkorange2", size = 0.25)
    }
}
print(Kenya.73.sq_vert_accel_ratio_speed.plot)
```

\pagebreak 

## Epoch 2: Display of Ratio betwwen Vertical Jolt and Speed

**sq_vert_jolt_ratio_speed**: the ratio between squared vertical acceleration (without natual gravity) and traveling speed \newline

```{r data cleaning 2, echo=TRUE}
# calculate ratio between vertical acceleration and speed
for (i in 1:nrow(Kenya.73)) {
    if (Kenya.73$Speed[i] == 0) {
        Kenya.73$sq_vert_jolt_ratio_speed[i] = 0
    }
    else {
        Kenya.73$sq_vert_jolt_ratio_speed[i] = (Kenya.73$vert_accel[i] * Kenya.73$vert_accel[i]) / Kenya.73$Speed[i]
    }
}
```

```{r exploratory data analysis 2.1, echo=FALSE}
require(ggplot2)
Kenya.73.sq_vert_jolt_ratio_speed.plot <- ggplot(Kenya.73[1:500, ], aes(x = time))
Kenya.73.sq_vert_jolt_ratio_speed.plot <- Kenya.73.sq_vert_jolt_ratio_speed.plot + geom_line(aes(y = sq_vert_jolt_ratio_speed))
Kenya.73.sq_vert_jolt_ratio_speed.plot <- Kenya.73.sq_vert_jolt_ratio_speed.plot + xlab("Time") 
Kenya.73.sq_vert_jolt_ratio_speed.plot <- Kenya.73.sq_vert_jolt_ratio_speed.plot + ylab("Ratio of Sq. Vert. Jolt and Speed") 
Kenya.73.sq_vert_jolt_ratio_speed.plot <- Kenya.73.sq_vert_jolt_ratio_speed.plot + ggtitle("Time-Indexed Ratio of Squared Vertical Jolt and Speed")
for (i in 1:500) {
    if (Kenya.73$label[i] == 1) {
         Kenya.73.sq_vert_jolt_ratio_speed.plot <- Kenya.73.sq_vert_jolt_ratio_speed.plot + 
             annotate("pointrange", x = Kenya.73$time[i], y = 0, ymin = 0, ymax = 0, colour = "red", size = 0.25)
    }
    if (Kenya.73$label[i] == 2) {
        Kenya.73.sq_vert_jolt_ratio_speed.plot <- Kenya.73.sq_vert_jolt_ratio_speed.plot + 
             annotate("pointrange", x = Kenya.73$time[i], y = 0, ymin = 0, ymax = 0, colour = "darkorange2", size = 0.25)
    }
}
print(Kenya.73.sq_vert_jolt_ratio_speed.plot)
```

\pagebreak 

```{r exploratory data analysis 2.2, echo=FALSE}
require(ggplot2)
Kenya.73.sq_vert_jolt_ratio_speed.plot <- ggplot(Kenya.73[501:1000, ], aes(x = time))
Kenya.73.sq_vert_jolt_ratio_speed.plot <- Kenya.73.sq_vert_jolt_ratio_speed.plot + geom_line(aes(y = sq_vert_jolt_ratio_speed))
Kenya.73.sq_vert_jolt_ratio_speed.plot <- Kenya.73.sq_vert_jolt_ratio_speed.plot + xlab("Time") 
Kenya.73.sq_vert_jolt_ratio_speed.plot <- Kenya.73.sq_vert_jolt_ratio_speed.plot + ylab("Ratio of Sq. Vert. Jolt and Speed") 
Kenya.73.sq_vert_jolt_ratio_speed.plot <- Kenya.73.sq_vert_jolt_ratio_speed.plot + ggtitle("Time-Indexed Ratio of Squared Vertical Jolt and Speed")
for (i in 501:1000) {
    if (Kenya.73$label[i] == 1) {
         Kenya.73.sq_vert_jolt_ratio_speed.plot <- Kenya.73.sq_vert_jolt_ratio_speed.plot + 
             annotate("pointrange", x = Kenya.73$time[i], y = 0, ymin = 0, ymax = 0, colour = "red", size = 0.25)
    }
    if (Kenya.73$label[i] == 2) {
        Kenya.73.sq_vert_jolt_ratio_speed.plot <- Kenya.73.sq_vert_jolt_ratio_speed.plot + 
             annotate("pointrange", x = Kenya.73$time[i], y = 0, ymin = 0, ymax = 0, colour = "darkorange2", size = 0.25)
    }
}
print(Kenya.73.sq_vert_jolt_ratio_speed.plot)
```

\pagebreak 

## Epoch 3: Display of 5-Sliding-Window Statistics of Vertical Jolt

**vert_jolt_mean**: 5-sliding-window mean of vertical jolt \newline 
**vert_jolt_sd**: 5-sliding-window standard deviation of vertical jolt \newline 
**vert_jolt_min**: 5-sliding-window minimum of vertical jolt \newline 
**vert_jolt_max**: 5-sliding-window maximum of vertical jolt \newline 

```{r data cleaning 3, echo=TRUE}
# calculate 5-sliding-window mean of vertical jolt
for (i in 3:nrow(Kenya.73)-2) {
    Kenya.73$vert_jolt_mean[i] = mean(c(Kenya.73$vert_jolt[i-2], 
                                        Kenya.73$vert_jolt[i-1], 
                                        Kenya.73$vert_jolt[i], 
                                        Kenya.73$vert_jolt[i+1], 
                                        Kenya.73$vert_jolt[i+2]))
}
Kenya.73$vert_jolt_mean[1] = 0
Kenya.73$vert_jolt_mean[2] = 0
Kenya.73$vert_jolt_mean[nrow(Kenya.73) - 1] = 0
Kenya.73$vert_jolt_mean[nrow(Kenya.73)] = 0
# calculate 5-sliding-window standard deviation of vertical jolt
for (i in 3:nrow(Kenya.73)-2) {
    Kenya.73$vert_jolt_sd[i] = sd(c(Kenya.73$vert_jolt[i-2], 
                                        Kenya.73$vert_jolt[i-1], 
                                        Kenya.73$vert_jolt[i], 
                                        Kenya.73$vert_jolt[i+1], 
                                        Kenya.73$vert_jolt[i+2]))
}
Kenya.73$vert_jolt_sd[1] = 0
Kenya.73$vert_jolt_sd[2] = 0
Kenya.73$vert_jolt_sd[nrow(Kenya.73) - 1] = 0
Kenya.73$vert_jolt_sd[nrow(Kenya.73)] = 0
# calculate 5-sliding-window minimum of vertical jolt
for (i in 3:nrow(Kenya.73)-2) {
    Kenya.73$vert_jolt_min[i] = min(c(Kenya.73$vert_jolt[i-2], 
                                        Kenya.73$vert_jolt[i-1], 
                                        Kenya.73$vert_jolt[i], 
                                        Kenya.73$vert_jolt[i+1], 
                                        Kenya.73$vert_jolt[i+2]))
}
Kenya.73$vert_jolt_min[1] = 0
Kenya.73$vert_jolt_min[2] = 0
Kenya.73$vert_jolt_min[nrow(Kenya.73) - 1] = 0
Kenya.73$vert_jolt_min[nrow(Kenya.73)] = 0
# calculate 5-sliding-window maxium of vertical jolt
for (i in 3:nrow(Kenya.73)-2) {
    Kenya.73$vert_jolt_max[i] = max(c(Kenya.73$vert_jolt[i-2], 
                                        Kenya.73$vert_jolt[i-1], 
                                        Kenya.73$vert_jolt[i], 
                                        Kenya.73$vert_jolt[i+1], 
                                        Kenya.73$vert_jolt[i+2]))
}
Kenya.73$vert_jolt_max[1] = 0
Kenya.73$vert_jolt_max[2] = 0
Kenya.73$vert_jolt_max[nrow(Kenya.73) - 1] = 0
Kenya.73$vert_jolt_max[nrow(Kenya.73)] = 0
```

```{r exploratory data analysis 3.1, echo=TRUE}
require(ggplot2)
Kenya.73.vert_jolt_5.plot <- ggplot(Kenya.73[1:500, ], aes(x = time))
Kenya.73.vert_jolt_5.plot <- Kenya.73.vert_jolt_5.plot + geom_line(aes(y = vert_jolt_mean, colour = "Mean"))
Kenya.73.vert_jolt_5.plot <- Kenya.73.vert_jolt_5.plot + geom_line(aes(y = vert_jolt_sd, colour = "Standard Deviation"))
Kenya.73.vert_jolt_5.plot <- Kenya.73.vert_jolt_5.plot + geom_line(aes(y = vert_jolt_min, colour = "Minimum"))
Kenya.73.vert_jolt_5.plot <- Kenya.73.vert_jolt_5.plot + geom_line(aes(y = vert_jolt_max, colour = "Maximum"))
Kenya.73.vert_jolt_5.plot <- Kenya.73.vert_jolt_5.plot + scale_colour_manual(values = c("darkorchid4", "blue4", "chartreuse4", "goldenrod3"))
Kenya.73.vert_jolt_5.plot <- Kenya.73.vert_jolt_5.plot + theme(legend.position = c(0.15, 0.8))
Kenya.73.vert_jolt_5.plot <- Kenya.73.vert_jolt_5.plot + xlab("Time") 
Kenya.73.vert_jolt_5.plot <- Kenya.73.vert_jolt_5.plot + ylab("5-Sliding-Window Stats. of Vertical Jolt") 
Kenya.73.vert_jolt_5.plot <- Kenya.73.vert_jolt_5.plot + ggtitle("5-Sliding-Window Statistics of Vertical Jolt")
for (i in 1:500) {
    if (Kenya.73$label[i] == 1) {
         Kenya.73.vert_jolt_5.plot <- Kenya.73.vert_jolt_5.plot + 
             annotate("pointrange", x = Kenya.73$time[i], y = 0, ymin = 0, ymax = 0, colour = "red", size = 0.25)
    }
    if (Kenya.73$label[i] == 2) {
        Kenya.73.vert_jolt_5.plot <- Kenya.73.vert_jolt_5.plot + 
             annotate("pointrange", x = Kenya.73$time[i], y = 0, ymin = 0, ymax = 0, colour = "darkorange2", size = 0.25)
    }
}
print(Kenya.73.vert_jolt_5.plot)
```

\pagebreak 

```{r exploratory data analysis 3.2, echo=TRUE}
require(ggplot2)
Kenya.73.vert_jolt_5.plot <- ggplot(Kenya.73[501:1000, ], aes(x = time))
Kenya.73.vert_jolt_5.plot <- Kenya.73.vert_jolt_5.plot + geom_line(aes(y = vert_jolt_mean, colour = "Mean"))
Kenya.73.vert_jolt_5.plot <- Kenya.73.vert_jolt_5.plot + geom_line(aes(y = vert_jolt_sd, colour = "Standard Deviation"))
Kenya.73.vert_jolt_5.plot <- Kenya.73.vert_jolt_5.plot + geom_line(aes(y = vert_jolt_min, colour = "Minimum"))
Kenya.73.vert_jolt_5.plot <- Kenya.73.vert_jolt_5.plot + geom_line(aes(y = vert_jolt_max, colour = "Maximum"))
Kenya.73.vert_jolt_5.plot <- Kenya.73.vert_jolt_5.plot + scale_colour_manual(values = c("darkorchid4", "blue4", "chartreuse4", "goldenrod3"))
Kenya.73.vert_jolt_5.plot <- Kenya.73.vert_jolt_5.plot + theme(legend.position = c(0.8, 0.925))
Kenya.73.vert_jolt_5.plot <- Kenya.73.vert_jolt_5.plot + xlab("Time") 
Kenya.73.vert_jolt_5.plot <- Kenya.73.vert_jolt_5.plot + ylab("5-Sliding-Window Stats. of Vertical Jolt") 
Kenya.73.vert_jolt_5.plot <- Kenya.73.vert_jolt_5.plot + ggtitle("5-Sliding-Window Statistics of Vertical Jolt")
for (i in 501:1000) {
    if (Kenya.73$label[i] == 1) {
         Kenya.73.vert_jolt_5.plot <- Kenya.73.vert_jolt_5.plot + 
             annotate("pointrange", x = Kenya.73$time[i], y = 0, ymin = 0, ymax = 0, colour = "red", size = 0.25)
    }
    if (Kenya.73$label[i] == 2) {
        Kenya.73.vert_jolt_5.plot <- Kenya.73.vert_jolt_5.plot + 
             annotate("pointrange", x = Kenya.73$time[i], y = 0, ymin = 0, ymax = 0, colour = "darkorange2", size = 0.25)
    }
}
print(Kenya.73.vert_jolt_5.plot)
```